\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\setcounter{page}{1}
I owe a big debt of gratitude to the numerous people who have helped me over the last few years with advice, encouragement, and direction as I worked on this thesis. First, I want to thank my main supervisor, Dr Jing Ying Wong, for providing structure to my research and helping me navigate the bureaucratic hurdles in academia and industry. I also want to extend my gratitude to my second supervisor, Prof. Andy Chan, with whom I had several insightful and helpful conversations regarding my work. Prof. Andy's work ethic has been an invaluable example to my overall growth as a researcher. I am also very grateful to Dr. Tomas Maul for agreeing to serve as my thesis's internal and second examiner. 

To my industry collaborators from MES Group, you have been the cornerstone of my financial support and exposure to the industry dynamic. You have constantly challenged me to be better, do better, and strive for nothing but the best. 

To my father Vincent, you not only introduced me to science and mathematics but also taught me to think creatively. I owe everything that I am to you. My siblings Susan, Maria, Joseph, Junior, Angela and Cynthia, uncle Dan, my mothers Mary and Betty - you know how grateful I am for all the encouragement you poured unto me when this journey became too hard to handle. Lastly, I want to thank my beautiful lady Sonia for her incredible, unwavering support and for always believing in me, even when I did not. Without your understanding, none of this would have been possible.
